Typed Functional Genetic Programming

### 2013-07-18 tex/thesis.tex
[tex]

@@

GP constructs the population of solution programs from 
two sets of building blocks; set $F$ of functions 
and set $T$ of constants and variables. 
In classical GP \textit{closure} of those 
building blocks is required.
This means that all functions in $F$ must accept all 
constants and variables from $T$ as its arguments and produce 
such values that are also acceptable by all functions in $F$ as arguments.
Main motivation behind this thesis is to design and implement a system that
eliminates this constraint. 
This can be achieved by utilization of \textit{types}.\\

Utilization of types in GP is not a new idea; 
it was dealt with in various previous works 
\cite{yu01,montana95,haynes96,olsson94}.
The contribution of this thesis can be seen 
in that it deals with this task using concepts from
theorie of typed $\lambda$-calculus and combinatory logic
and shows how they can be used for benefit in GP.\\

Notable highlights are: 
\begin{itemize}
 \item \textit{Inhabitation tree} (a slight modification 
   of \textit{inhabitation machine} presented in \cite{barendregt10}
   suited better for implementation purposes) 
   is presented and used for \textit{generating of individuals}.
 \item \textit{Abstraction elimination} is utilized
   to enable use of simple \textit{mutation operator}.
 \item \textit{Polymorphic} generalization of inhabitation tree,
   enabling use of polymorphic functions as building blocks, 
   is presented.

@@

Design of a GP system utilizing these methods is presented.
This design aims to be generalization of classical GP 
developed by Koza \cite{koza92}, rather than being 
specialized variant of GP using more recent techniques.
This decision is based on belief that modern specialized 
techniques can be incorporated later and on belief
that simplicity of design of classical GP plays nicely
with testing those $\lambda$-calculus concepts in 
the evolutionary context.

The implementation aims

~\\[3em]Témata co zmínit v úvodu :\\

& Co je to GP, 
 && že je to část EVA
 && že v clasickym GP je closure
 && a ruzný prace ukazujou jak se ho zbavit

& Co je to typ
  && typ jedobře zaběhlá programátorská konstrukce
  && matematicky uchopuje typy a programy v teorii typovaného lambda kalkulu

& cílem této práce je 
  && ukázat jak spolu hezky souvysej
     &&& např skrzestavební bloky a context
  && a jak využít konstrukty typový teorie k prospěchu v GP
     &&& such as AE, beta,etha reduction,...
  && na základě takto zmapované podobnosti navrhnout systém a implementovat
     rather then build yet another GP system, generalize ten klasickej

& Začít asi tim že evoluce je cool a že
  tohle bude o tom jak do GP dát typy.
& Říct jak jde text:
  && Začnem popsanim klasickýho GP podle Kozy
  && Vyslovíme definice okolo lambda termu a typu
  && Vyslovíme definici inhabitačního stromu
     kterej je stromovou verzí IM od Barendrechta.
  && Matematicky popíšem jak na to dem my.
  && Popíšem rozšíření zakladního systému na 
     polymorfní.
  && Popíšem to z programátorskýho hlediska
  && Popíšem příklady na kterejch to demonstrujem
  && Srovnáme to s nějakejma přístupama co už 
     existujou 
& Možná se zmínit o tý představě dvou pólu
  induktivním a deduktivním -
  klasickym GP a Dokazovačem, čim složitější
  typovej sytem tim víc se jedná o deduktivní   
  systém a blíží se to dokazovači.
   && Ve světle toho, je pak hezky vidět, že
      je chytrý exhaustivně systematicky prohledávat od nejmenšího
      po největší- jak budeme zesložiťovat typovej systém, tak
      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
      jich generovat kvanta
& Nějak naťuknout už tady v úvodu jak se to má při
  srovnání s tim jak na to dou jiný (Yu atd)
& Výkřiky co zmínit:
  && Na co zaměřený:
     &&& Proof of a concept
     &&& Pořádně to vymáchat v typový teorii
     &&& Spíš než bejt další z řady nástroju
         se snaží zobecnovat ty už existující
     &&& Spíš jednoduchost
         &&&& Ve většině aspektú to sleduje
              Kozu - tzn i v málo řídících
              parametru
     &&& Duraz na generování termu (!)              
     &&& Po implementační stránce
         &&&& Sranda s JS překladem
         &&&& Servrová architektura
         &&&& oboje spíš srandy
  && Na co nezaměřený (ale snad taky ok v tom):
     &&& Performance   
     &&& Mature knihovna co by šla použít robustně
         v cizim kódu - spíš to počítá že to 
         bude dál rozvíjeno a tohle je mezizastávka
\end{easylist}

@@

\textit{Genetic programming} is a technique inspired by biological evolution
that for a given problem tries to find computer programs able to solve that problem. 
GP was developed by John Koza \cite{koza92} in 1992.

@@

Another crucial input besides fitness function is a collection of \textit{building blocks}.
It is collection of symbols (accompanied with an information about number of arguments).
Those symbols are used to construct trees representing individuals.  
\\\\
\Lets describe GP algorithm briefly:

@@

This step consists of two sub steps\footnote{TODO : Technically it is done in a little  
bit different fashion which is equivalent.}:
\begin{itemize} 
	\item Selection of \textit{parents} for individuals of the next population based on the fitness.
	      The bigger fitness of an individual of the current population is, 
	      the better chance of success being selected as parent it has.  
	\item Application of genetic operators (such as \textit{crossover}, 
	      \textit{reproduction} and \textit{mutation}) 
		  on parent individuals producing new individuals of the next population.  
\end{itemize}	  
This transformation is repeatedly applied for a predefined number of steps (which is called 
number of \textit{generations}) or until some predefined criterion is met.	
\\\\
\Lets now look on GP at more detail.

@@

In GP programs are represented as expressions. \Lets define expression inductively:

\begin{todo}
 & Asi to přejmenvat na S-expression.  
\end{todo}

\begin{itemize}
	\item Constant\footnote{By constants we also mean procedures with zero  	
		  arguments.} or variable symbol $s$ is expression.
	\item Let there be a function with symbol $f$ which has $n$ arguments. 
	      And let there be expressions $e_{1}, ..., e_{n}$. 
	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression
	      \footnote{This notation comes from Lisp programming language, 
	      classical notation would be $f(e_{1}, ... ,e_{n})$. }.   
\end{itemize}

@@

\begin{itemize}
	\item One node tree $s$.
    \item \Tree
			[.$f$	
		 		\text{$t_{1}$}
		 		\text{...}
		 		\text{$t_{n}$} ]\\\\
		 Where $t_{1}, ..., t_{n}$ are trees corresponding to expressions $e_{1}, ..., e_{n}$.	   
\end{itemize}

@@

\begin{itemize}
	\item Terminal set $T$ : Set of symbols used as leaf nodes of 	               
	      program trees standing for constants and variables.
	\item Function set $F$ : Set of symbols used as internal nodes 
	      of program trees standing for functions.
\end{itemize}

@@

Beside symbols in \TuF there must also be 
an implementation for each symbol which is not variable. 
And for every function symbol from $F$ there must be specified 
number of arguments.\\

@@

\begin{todo}
& koza tomu řiká closure, přečíst si tu pasáž (str 81) a 
upravit podle toho tady tu terminologii
\end{todo}

Satisfying this constrain ensures that every program tree build 
from \TuF will be total.

We will refer to this constrain as \textit{over-one-type} constrain
in the further text.\\

We can say that the motivation behind this work is to construct system where we
eliminate this constraint.

@@

We will describe three tree generating method described by Koza 
in \cite{koza92} called \textit{ramped half-and-half}. \\

@@

There are two generating sub-methods called \textit{full} and 
\textit{grow} each restricting differently the subset of \TuF. 
This restriction depends on depth of the
node for which the symbol is being selected.

In order to perform \textit{ramped half-and-half} method
one of those two sub-methods is randomly (uniformly) 
selected and the selected one is performed.
This selection is done for each generated tree, so there
is 50\% chance to be generated by \textit{full} 
and 50\% chance to be generated by \textit{grow}.\\

A maximum depth $D_{initial}$ is defined (e.g. 6).

@@

In \EA there is plenty of options for selection mechanisms 
for us to choose from. Again we will describe mechanism Koza
used in his first book on GP. It is the \textit{Roulette selection}.
It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\

@@

For Koza the most important genetic operator in GP is 
crossover. It is operator inspired by sexual reproduction
occurring in the nature. Generally speaking crossover takes
two (parent) individuals and combines theirs genomes to produce 
two possibly new (child) individuals.

In GP the most common crossover is \textit{Subtree swapping crossover}.
This crossover randomly selects one node in each parent tree.
Two new child individuals are constructed by swapping subtrees 
which have roots in those selected nodes.\\

@@

\begin{todo}
 & zmínit maximální hloubku vzniklýho dítěte (17)
\end{todo}

@@

@@

Let $pop_{t}$ be the current population which we want to 
transform into the next population $pop_{t+1}$. 
We start by initializing $pop_{t+1}$ by empty population.

@@

@@

\begin{todo}
 & mention best preservation and same popSize checking.
 & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
   defining EA and then specifying the differences or something 
   like that...    
 & History, citations, etc ....
 & ten algoritmus v kozovi dělá "Designate Result" já tam vracim 
   poslední populaci
\end{todo}

@@

\begin{todo}
 & Shrnout tu nějak ve skratce tuhle kapitolu.. 
\end{todo}

\section{Lambda term}
\label{deflam}

@@

Let $V$ be infinite countable set of {\it 
variable names}.  \\* 
Let $C$ be set of {\it constant names}.	 \\*		
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~~($\lambda$-abstraction)} 
\end{align*}

Function application and $\lambda$-abstraction are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}. \\

@@

We use bold upper case symbols such as 
$\Scomb, \Kcomb, \Icomb$ as abbreviation
for specific terms, e.i. 
$\Icomb = \lamb{x}{x}$.  
\\

@@

\begin{todo}
	& říct co to vubec je ty intutivně (v par větách jen) 
	  a nějak to napojit na to jak to používám já
	  pak v tom GP.
	& Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
\end{todo}

\section{Type}
\label{deftype}

@@

Let $A$ be set of {\it atomic type names}. \\*
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows:

\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$. \\

We use following notation
abbreviation for better readability:\\

@@

\begin{todo}
   & určitě promluvit nějak v úvodu k tomu že se jedná o 
	 Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
   & říct že se vždycky zastavěj (jak se to správně jmenuje?) ale
     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
     říct to někde kde už se ví jak sou napojený na termy    
   & možná se hlobějc zmínit o curryingu  
\end{todo}

@@

@@

\begin{enum}
 & It should eliminate the \textit{over-one-type} constrain of classical GP
   (see \ref{building-blocks}). 
 & It should be generalization of the classical GP described in \ref{GP},
   by which is meant that for \textit{building blocks} satisfying 
   constrains of classical GP there should be simple setup of control parameters
   of the system which will make the system behave in precisely the same way
   as the classical GP behaves. 
 & It should utilize the theory around typed $\lambda$-calculus.  
\end{enum}

@@

\subsection{Using term generation}

@@

\textbf{Poznámky :\\
bylo by asi mimojiný dobrý 
ospravedlnit/okomentovat tak fancy (=javascript frontend) 
implementaci při 
dost nedodělanejch jinejch věcech (heuristika by 
mohla být lepší, mutace by mohli bejt, atd)
ale věřil sem, že je to proces iterativní
a že je po celou dobu dobrý mít na čem to testovat.
Javascript byl vybranej pač HTML je nejpromáklejší
UI jaký vubec je na zemi teď (state of dzí art)
a tak to snad nebyla stráta času.
navíc se naskitla díky tomu další pěkná příležitost
k ukázání toho, že to neni omezený na haskell
ale že se kombinatorový konstrukty daj v zásadě přeložit do libovolnýho jazyka celkem bez problémů
nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
spíš nějaky podrobnosti přehodit do implementačních detailu
a tady to jen tak líznout.}

@@

@@

>>>>

% dát pryč will, používat přítomnej čas

@@

GP constructs the population of solution programs from 
two sets of building blocks; set $F$ of functions 
and set $T$ of constants and variables. 
In classical GP \textit{closure} of those 
building blocks is required.
This means that all functions in $F$ must accept all 
constants and variables from $T$ as their arguments and produce 
such values that are also acceptable by all functions in $F$ as arguments.
Main motivation behind this thesis is to design and implement a system that
eliminates this constraint. 
This can be achieved by utilization of \textit{types}.\\

Utilization of types in GP is not a new idea; 
it was dealt with in various previous works 
\cite{yu01,montana95,haynes96,olsson94}.
The contribution of this thesis can be seen 
in that it deals with this task using concepts from
theory of typed $\lambda$-calculus and combinatory logic
and shows how to use them in GP.
Hopefully, it is also shown that those two fields are
nicely linked together, e.g., the concept of \textit{building blocks}
corresponds to concept of \textit{context}. \\

Notable highlights are: 
\begin{itemize}
 \item \textit{Inhabitation tree} (a slight modification 
   of \textit{inhabitation machine} presented in \cite{barendregt10}
   suited better for reasoning about implementation details) 
   is presented and used for \textit{generation of individuals}.
   This generation process is parameterized by simple 
   \textit{search strategy}. Several search strategies are presented.   
 \item \textit{Abstraction elimination} is utilized
   to enable use of simple \textit{mutation operator}.
 \item \textit{Polymorphic} generalization of inhabitation tree,
   enabling use of polymorphic functions as building blocks, 
   is presented.

@@

Design of a GP system utilizing these methods is presented.
This design aims to be generalization of classical GP 
developed by Koza \cite{koza92}, rather than being 
specialized variant of GP using more recent techniques.
This decision is based on belief
that simplicity of design of classical GP plays nicely
with testing those $\lambda$-calculus concepts in 
the evolutionary context and on belief that modern specialized 
techniques can be incorporated later.

%The implementation of our system aims 
Described system is implemented in purely functional language 
\textit{Haskell}. Evaluation of individual \lterms is 
performed by translating them to
Haskell programs and evaluating them in Haskell interpreter. 
The core part of the system runs as server, which is 
managed through user interface accessible via
web browser. Solution individuals are also translated
to \textit{JavaScript} expressions, which enables  
an interactive solution presentation.
This approach helps us with demonstrating 
that \lterms can be easily translated to a 
widely used programming language.

Several example problems for the designed system are presented 
and examined. Those examples are not intended to 
bring breakthrough results; they are intended to 
serve as demonstration of that the system is working properly
and as demonstration of several directions in which the system
can be used.

\begin{todo}
 & možná přidat ještě něco z poznámek co jsou teď v Conclusion
\end{todo}

@@

\textit{Genetic programming} is a technique inspired by biological evolution
that for a given problem tries to find computer programs able to solve that problem. 
GP was developed by John Koza in 1992 \cite{koza92}.

@@

Another crucial input besides fitness function is a collection of \textit{building blocks}.
It is collection of symbols (accompanied with an information about number of arguments).
Those symbols are used to construct trees representing individuals.\\

\Lets describe GP algorithm briefly:

@@

This step consists of two sub steps:
\begin{itemize} 
	\item Selection of \textit{parents} for individuals of the next population based on the fitness.
	      The bigger fitness of an individual of the current population is, 
	      the better chance of success being selected as parent it has.  
	\item Application of genetic operators (such as \textit{crossover}, 
	      \textit{reproduction} and \textit{mutation}) 
		  on parent individuals producing new individuals of the next population.  
\end{itemize}	  
This transformation is repeatedly applied for a predefined number of steps (which is called 
number of \textit{generations}) or until some predefined criterion is met.	
\\\\
\Lets now look on GP at more detail.

@@

In GP programs are represented as \textit{S-expressions}. 
\Lets define S-expression inductively:

\begin{samepage}
\begin{itemize}
	\item Constant or variable symbol $s$ is S-expression.\footnote{
	      By constants we also mean procedures with zero arguments.}
	\item Let there be a function with symbol $f$ which has $n$ arguments.

	      And let there be S-expressions $e_{1}, ..., e_{n}$.

	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression.\footnote{
	      This notation comes from Lisp programming language, 
	      a more standard notation would be $f(e_{1}, ... ,e_{n})$.}   
\end{itemize}
\end{samepage}

@@

\begin{itemize}
	\item One node tree $s$.
    \item \Tree
			[.$f$	
		 		\text{$t_{1}$}
		 		\text{...}
		 		\text{$t_{n}$} ]\\\\
		 Where $t_{1}, ..., t_{n}$ are trees corresponding 
		 to S-expressions $e_{1}, ..., e_{n}$.	   
\end{itemize}

@@

\begin{itemize}
	\item \textit{Terminal set $T$}: Set of symbols used as leaf nodes of 	               
	      program trees standing for constants and variables.
	\item \textit{Function set $F$}: Set of symbols used as internal nodes 
	      of program trees standing for functions.
\end{itemize}

@@

For each symbol in \TuF that is not variable 
there must also be an implementation.
And for every function symbol from $F$ there must be specified 
number of arguments.\\

@@

Satisfying this constrain ensures that every program tree build 
from \TuF will be total. We will refer to this constrain as to \textit{closure}.\\

We can say that the motivation behind this thesis is to construct system,
where we eliminate this constraint.

@@

We will describe tree generating method described by Koza 
called \textit{ramped half-and-half}  \cite{koza92}. \\

@@

There are two generating sub-methods called \textit{full} and 
\textit{grow}; each of them restricting differently the subset of \TuF. 
This restriction depends on depth of the
node for which the symbol is being selected.

In order to perform \textit{ramped half-and-half} method
one of those two sub-methods is randomly (uniformly) 
selected and the selected one is performed.
This selection is done for each generated tree, so there
is 50\% chance for tree to be generated by \textit{full} 
and 50\% chance for tree to be generated by \textit{grow}.\\

A maximum depth $D_{initial}$ is defined (e.g. $D_{initial} = 6$).

@@

In the field of evolutionary algorithms there is plenty of 
options for selection mechanisms 
for us to choose from. Again we will describe mechanism Koza
used in his first book on GP. It is the \textit{Roulette selection}.
It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\

@@

For Koza the most important genetic operator in GP is 
crossover. It is operator inspired by sexual reproduction
occurring in the nature. Generally speaking crossover takes
two (parent) individuals and combines theirs genomes to produce 
two possibly new (offspring) individuals.

In GP the most common crossover is \textit{Subtree swapping crossover}.
This crossover randomly selects one node in each parent tree.
Two new child individuals are constructed by swapping subtrees 
that have roots in those selected nodes.\\

@@

A maximum permissible depth $D_{created}$ 
for offspring individuals is defined (e.g. $D_{created} = 17$).
If one of the offspring has greater depth than this limit, then 
this offspring is replaced by the first parent in the result of 
the crossover operator. If both offspring exceeds this limit, than 
both are replaced by both parents.

@@

Similarly as for crossover, a maximum permissible depth for mutant individual 
is defined.

@@

Let $pop_{t}$ be the current population that we want to 
transform into the next population $pop_{t+1}$. 
We start by initializing $pop_{t+1}$ by empty population.

@@

\section{Evaluation}

\begin{todo}
 & !!!
 & zmínit tu terminate bool
\end{todo}

@@

\begin{todo}
 & mention best preservation and same popSize checking.
 & popsat co je to popWithF
 & popsat co je to terminate 
 & ten algoritmus v kozovi dělá "Designate Result" já tam vracim 
   poslední populaci
\end{todo}

@@

In this chapter we briefly describe key concepts
from simply typed $\lambda$-calculus.
We try to favor readability over formality,
for more detailed and formal treatment of this subject see 
\cite{barendregt92}.

After that, \textit{inhabitation tree} is introduced.
It is structure used for reasoning about generation of individuals.

\theoremstyle{plain} 
\newtheorem{theorem}{Theorem} 
\newtheorem{proposition}{Proposition} 
\newtheorem{lemma}{Lemma} 
\newtheorem*{corollary}{Corollary}

\theoremstyle{definition} 
\newtheorem*{definition}{Definition} 
\newtheorem{conjecture}{Conjecture}
 \newtheorem*{example}{Example}

\theoremstyle{remark} 
\newtheorem*{remark}{Remark} 
\newtheorem*{note}{Note} 
\newtheorem{case}{Case}

\section{\lterms}
\label{deflam}

Now we will formally describe programming language 
in which we will generate individual programs --- so called \lterms.

@@

\begin{definition}

Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~~($\lambda$-abstraction)} 
\end{align*}~
\end{definition}

\textit{Function application} and 
\textit{$\lambda$-abstraction} are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}. \\

@@

We use bold upper case symbols such as 
$\Scomb, \Kcomb, \Icomb$ as abbreviation
for specific terms, e.i. 
$\Icomb = \lamb{x}{x}$.  
\\[1em]

@@

\section{Types}
\label{deftype}

@@

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).
After putting those two pieces 
(\textit{\lterms} and \textit{types}) together 
we will get system called \textit{simply typed $\lambda$-calculus}.

\begin{definition}
Let $A$ be set of {\it atomic type names}. \\
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}~

\end{definition}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$.

Usually A is called \textit{set of type variables}, 
but since it might be source of confusion with polymorphic type variables
we prefer this name, which also better suits our usage. 
\\

We use following notation
abbreviation for better readability:

@@

\begin{todo}  
   & možná se hlobějc zmínit o curryingu  
\end{todo}

@@

\section{Simply typed $\lambda$-calculus}

\begin{todo}
   & říct že se vždycky zastavěj (strong normalization) ale
     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
     říct to někde kde už se ví jak sou napojený na termy    
\end{todo}

@@

\begin{enum}
 & It should eliminate the \textit{closure} constrain of classical GP
   (see \ref{building-blocks}). 
 & It should be generalization of the classical GP described in \ref{GP},
   by which is meant that for \textit{building blocks} satisfying 
   constrains of classical GP there should be simple setup of control parameters
   of the system which will make the system behave in precisely the same way
   as the classical GP behaves. 
 & It should utilize the theory around typed $\lambda$-calculus.  
\end{enum}

@@

%\subsection{Using term generation}

@@

\begin{todo}
 & !!!!!!!!!!!!!!!!!
\end{todo}

Další témata co zmínit v úvodu nebo v závěru :\\

& Možná se zmínit o tý představě dvou pólu
  induktivním a deduktivním -
  klasickym GP a Dokazovačem, čim složitější
  typovej sytem tim víc se jedná o deduktivní   
  systém a blíží se to dokazovači.
   && Ve světle toho, je pak hezky vidět, že
      je chytrý exhaustivně systematicky prohledávat od nejmenšího
      po největší- jak budeme zesložiťovat typovej systém, tak
      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
      jich generovat kvanta

& že imlementace je delana těma vtipnejamtajpklasama umožnujicima libovolný
  evoluční struktury na evolvovani

& říct že vnekterejch aspektech je videt, že to je work in progress
  ale že si myaslime že i tak to ukazuje pěkný výsledky 
  a že doufame že ty nedodělky budou další prací odstraněny

& Říct jak jde text:
  && Začnem popsanim klasickýho GP podle Kozy
  && Vyslovíme definice okolo lambda termu a typu
  && Vyslovíme definici inhabitačního stromu
     kterej je stromovou verzí IM od Barendrechta.
  && Matematicky popíšem jak na to dem my.
  && Popíšem rozšíření zakladního systému na 
     polymorfní.
  && Popíšem to z programátorskýho hlediska
  && Popíšem příklady na kterejch to demonstrujem
  && Srovnáme to s nějakejma přístupama co už 
     existujou

& Nějak naťuknout už tady v úvodu jak se to má při
  srovnání s tim jak na to dou jiný (Yu atd)

& Výkřiky co zmínit:
  && Na co zaměřený:
     &&& Proof of a concept
     &&& Spíš jednoduchost
         &&&& Ve většině aspektú to sleduje
              Kozu - tzn i v málo řídících
              parametru
     &&& Duraz na generování termu (!)

  && Na co nezaměřený (ale snad taky ok v tom):
     &&& Performance   
     &&& Mature knihovna co by šla použít robustně
         v cizim kódu - spíš to počítá že to 
         bude dál rozvíjeno a tohle je mezizastávka

 & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
   defining EA and then specifying the differences or something 
   like that...

& bylo by asi mimojiný dobrý 
ospravedlnit/okomentovat tak fancy (=javascript frontend) 
implementaci při 
dost nedodělanejch jinejch věcech (heuristika by 
mohla být lepší, mutace by mohli bejt, atd)
ale věřil sem, že je to proces iterativní
a že je po celou dobu dobrý mít na čem to testovat.
Javascript byl vybranej pač HTML je nejpromáklejší
UI jaký vubec je na zemi teď (state of dzí art)
a tak to snad nebyla stráta času.
navíc se naskitla díky tomu další pěkná příležitost
k ukázání toho, že to neni omezený na haskell
ale že se kombinatorový konstrukty daj 
v zásadě přeložit do libovolnýho jazyka celkem bez problému
nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
spíš nějaky podrobnosti přehodit do implementačních detailu
a tady to jen tak líznout.

\end{easylist}

@@

\bibitem{barendregt92}
H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum , H.  P.
\emph{Lambda Calculi with Types.} 
Handbook of Logic in Computer Science, 1992.

@@

\bibitem{jones87}
Simon Peyton Jones. 
\emph{The Implementation of Functional Programming Languages}. 
Prentice Hall, 1987.
